Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex
text2

The models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e., to represent the plant dynamics $A_d$, $B_d$ and its states. 
The second is the quantization error introduced by the digital controller, 
which performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by fixed-point representation of numbers.

Let $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.

The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}_{\langle I,F \rangle}(x)$ be the same number represented in a fixed point domain as above. $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands. 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.

\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item In this implementation we only need to divide by positive numbers(where it is needed for checking stability). 
\end{enumerate}

@@

We use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model. Thus all mathematical operations occurring in the digital controller occur in the domain $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations are carried out in the domain $\mathcal{F}_{\langle I_p,F_p \rangle}$.

We will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model ($I_p \geq I_c \wedge F_p \geq F_c$). Thus any mathematical operations occurring in the digital controller will occur in the range of $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations will be carried out in the range of $\mathcal{F}_{\langle I_p,F_p \rangle}$.

>>>>

The models we consider have two sources of error due to numeric representation. 
The first is the numeric error introduced by the fixed-point numbers employed to model the plant, i.e., to represent the plant dynamics $A_d$, $B_d$ and its states. 
The second is the quantization error introduced by the digital controller, 
which performs operations in the domain of fixed-point numbers. 
In this section we formally outline the errors introduced by the fixed-point representation of numbers. 
% 
Let $\mathcal{F}_{\langle I,F \rangle}(x)$ denote a real number $x$ represented in a fixed point domain, with $I$ bits representing the integer part and $F$ bits representing the decimal part. The smallest number that can be represented in this domain is $c_m=2^{-F}$.

The following approximation errors will arise:
\begin{enumerate}
\item {\bf Truncation:}
Let $x$ be a real number, and $\mathcal{F}_{\langle I,F \rangle}(x)$ be the same number represented in a fixed-point domain as above. 
Then $\mathcal{F}_{\langle I,F \rangle}(x) = x-\delta_T : \delta_T=x\ \%_{c_m}\ \tilde x$, where $\%_{c_m}$ is the modulus operation performed on the last bit.

So $\delta_T$ is the truncation error and it will propagate across operations.
\item {\bf Rounding:}
The following errors appear in simple operations. Let $c_1, c_2$ and $c_3$ be real numbers, and $\delta_{T1}$ and $\delta_{T2}$ be the truncation errors caused by representing $c_1$ and $c_2$ in the fixed-point domain as above.
\begin{enumerate}
\item Addition/Subtraction: these operations only propagate errors coming from truncation of the operands, namely 
$\mathcal{F}_{\langle I,F \rangle}(c_1) \pm \mathcal{F}_{\langle I,F \rangle}(c_2) = c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}| + |\delta_{T2}|$.

\item Multiplication: $\mathcal{F}_{\langle I,F \rangle}(c_1) * \mathcal{F}_{\langle I,F \rangle}(c_2) =  c_3 + \delta_3 :|\delta_3| \leq |\delta_{T1}*\mathcal{F}_{\langle I,F \rangle}(c_2)| + |\delta_{T2}*\mathcal{F}_{\langle I,F \rangle}(c_1)| + c_m$, where $c_m=2^{-F}$ as above.
\item In this implementation we only need to divide by positive numbers(where it is needed for checking stability) \todo{so: what does that imply? pls clarify explicitly.}. 
\end{enumerate}

@@

We will use $\mathcal{F}_{\langle I_c,F_c \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the controller, and $\mathcal{F}_{\langle I_p,F_p \rangle}(x)$ to denote a real number $x$ represented in the fixed-point precision level of the plant model ($I_p \geq I_c \wedge F_p \geq F_c$). 
\todo{Should we include a discussion on the fact that the plant is physical? So for example, a reviewer may ask what is the meaning of the model with time derivative, with variables over a Euclidean space. }
Thus any mathematical operations occurring in the digital controller will occur in the range of $\mathcal{F}_{\langle I_c,F_c \rangle}$, and all other calculations will be carried out in the range of $\mathcal{F}_{\langle I_p,F_p \rangle}$.
